On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-08-15 paper.tex
Writing RS-theorem

The well-known counterexample of P.~Cohn shows that the $\K_1$-analogue of Serre problem fails for $\K_1(2, -)$ and thus Suslin's result cannot be improved, see~\cite[\S~I.8]{Lam10}
Similarly, it is known that an analogue of Serre problem fails for $\K_2(3, -)$, see~\cite{We14}.
Thus, our~\cref{thm:horrocks-sl4} completely settles the remaining case $n=4$ not already covered by the results of Wendt and Tulenbayev.

Previously in~\cite{LS17} A.~Lavrenov and the author showed that the local-global principle holds for $\St(4, -)$.
As said before, this together with~\cref{thm:horrocks-sl4} allows us to show the following.
\begin{thm} \label{thm:Serre-problem} %TODO: Fix statement
  For any field $k$ there is an isomorphism $\K_2(4, k[t_1, \ldots, t_n]) \cong \K_2(4, k)$.
  In fact Tulenbayev early stability theorem can be proved starting with the bound $n=4$... 
\end{thm}

@@

\subsection{Rehmann-Soul{\'e} theorem}

@@

For $n \geq 1$ consider the following collections of homogeneous generators:
\[\XX{n} = \{ x_\alpha(\xi) \mid \xi \in B_k,\text{ for } k\leq n\}. \]
It is obvious that $\XX{n} \subseteq \XX{n+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

Let $\Phi$ be a simply-laced irreducible root system of rank $\geq 3$.
For $m\geq 1$ denote by $G_m$ the group presented by generators $\mathcal{X}_m$ and the set of relations $\RR{m}$ consisting of the following
 three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_\alpha(\xi) x_\alpha(\eta) & = x_\alpha(\xi+\eta),&  \xi,\eta\in B_k,\ k\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_k,\ \eta \in B_l,\ l,k\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_k,\ \eta \in B_l,\ l, k, l+k \leq m.
 \end{align}

@@

%\begin{rem}
%There are two reasons why we can't simply refer to~\cite{Re75} or~\cite{RS76} for the proof of~\cref{lem:tul3.3}.
%The first is that we want to prove Rehmann and Soul{\'e} prove that $G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$, while we want to %establish this fact starting from $m = 1$.
%The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%(it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}).
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).
%\end{rem}

>>>>

The well-known counterexample of P.~Cohn shows that the $\K_1$-analogue of Serre problem fails for $\K_1(2, -)$ which shows that Suslin's result cannot be improved, see~\cite[\S~I.8]{Lam10}
Similarly, it is known that an analogue of Serre problem fails for $\K_2(3, -)$, see~\cite{We14}.
Thus, our~\cref{thm:horrocks-sl4} completely settles the remaining case $n=4$ not already covered by the results of Wendt and Tulenbayev.

Previously in~\cite{LS17} A.~Lavrenov and the author showed that the local-global principle holds for $\St(4, -)$.
As said before, this together with~\cref{thm:horrocks-sl4} allows us to show the following result, which is an improvement of Tulenbayev stability theorem.
\begin{thm} \label{thm:Serre-problem} %TODO: Fix statement
  For any field $k$ there is an isomorphism $\K_2(4, k[t_1, \ldots, t_n]) \cong \K_2(4, k)$.
  In fact Tulenbayev early stability theorem can be proved starting with the bound $n=4$... 
\end{thm}
It is clear that the positive solution of Serre problem for $\K_2(n, -)$ for $n\geq 4$ is a corollary of this result.
%TODO: Add couple of further corollaries

@@

\section{Rehmann--Soul{\'e} theorem}
The aim of this section is to prove a variant of Rehmann--Soul{\'e} theorem on presentation of Steinberg groups.

@@

In what follows we will be concerned with finding a compact presentation of the Steinberg group $\St(4, B)$. This presentation is analogous to the that~\cite[Lemma~3.3]{Tu83}, which is formulated for $\St(n, B)$, $n\geq 5$.

For $m \geq 1$ consider the following set of generators:
\[\XX{m} = \{ x_{ij}(\xi) \mid \xi \in B_k,\text{ for } k\leq m,\ 1\leq i, j\leq 4,\ i\neq j\}. \]
It is obvious that $\XX{m} \subseteq \XX{m+1}$, denote by $\XX{\infty}$ the union of all $\XX{n}$'s.

For $m\geq 1$ denote by $G_m$ the group presented by generators $\mathcal{X}_m$ and the set of relations $\RR{m}$ consisting of the following three families of relations:
\begin{align}
 \label{eq:am} \tag{$a_m$} x_{ij}(\xi) x_{ij}(\eta) & = x_{ij}(\xi+\eta),&  \xi,\eta\in B_k,\ k\leq m;&\\
 \label{eq:bm} \tag{$b_m$} [x_\alpha(\xi), x_{\alpha'}(\eta))] &  = 1, & \text{in the case $\alpha+\alpha'\not\in\Phi\cup\{0\}$,}\\
 \nonumber                                                     &       & \xi \in B_k,\ \eta \in B_l,\ l,k\leq m;\\
 \label{eq:cm} \tag{$c_m$} [x_\alpha(\xi), x_{\alpha'}(\eta)] & = x_{\alpha+\alpha'}(N_{\alpha,\alpha'}\xi\eta), & \text{in the case $\alpha+\alpha'\in \Phi$,}\\
 \nonumber                                                    &  & \xi \in B_k,\ \eta \in B_l,\ l, k, l+k \leq m.
 \end{align}

There are several reasons why we can't simply refer to the original results of U.~Rehmann and C.~Soul{\'e}~\cite{Re75} or~\cite{RS76}.
The first is that we want to prove Rehmann and Soul{\'e} prove that $G_m \to G_{m+1}$ is an isomorphism for $m\geq 2$, while we want to %establish this fact starting from $m = 1$.
%The second is that they prove their results only in the special case $B = A[t]$, moreover, they make assumptions about the ground ring $A$
%(it is assumed that $A=k$ is a field in~\cite{Re75} and that $A=\mathbb{Z}$ in~\cite{RS76}).
% (this can be replaced by the weaker assumption that $A$ is additively generated by its units).

@@
